---
title: "Beer and Brew Project"
author: "Xavier Mojica"
date: "2022-10-15"
output: 
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Beer and Brewery Data
## Presenting Findings of Beer and Breweries 
### R Data Manipuilationn and Import. Pre load packages
```{r}
library(knitr)
library(e1071)
library(caret)
library(class)
library(tidyverse)
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(rmarkdown)
```

###Read in file
```{r}
#Reading our Beer and Brewery Files
totalBeer = read.csv('~/Dropbox/Mac/Desktop/Desktop/Intro-DataScience/Unit7/Unit8and9/Beers.csv', header = TRUE)
totalBrew = read.csv('~/Dropbox/Mac/Desktop/Desktop/Intro-DataScience/Unit7/Unit8and9/Breweries.csv', header = TRUE)
```
###Define a graph with breweries per state
```{r}
#1) How many breweries are present in each state?
totalBrew %>% group_by(State) %>% count(State)
ggplot(totalBrew, aes(x = State), fill = State) + geom_bar() + labs(title ="Amount of Breweries per State", x = "States", y = "Amount of Breweries")
```
##The following states are shown with each state and the amount of breweries it has.Alaska has 7, Alabama ahs 3, Arkansas has 2, Arizona has 11, California has 39, Colorado has 47 and so forth. The rest of the list may be found on the HTML document. 


###Merge both data
```{r}
#2) Merging Beer and Brewery data
#Checking dimensions of both dataframes 
dim(totalBeer)
dim(totalBrew)

#Merging Beer is totalBeer and Brewery is totalBrew.
mergedB = merge(totalBeer, totalBrew, by.x = "Brewery_id", by.y = "Brew_ID")
dim(mergedB)

#Looking at the first 6 observations.
head(mergedB, n=6)

#Looking at the last 6 observations.
tail(mergedB, n = 6)
```
###Cleaning data
```{r}
#3)	Addressing the missing values in each column.
##Cleaning up data by removing NAs. 
mergedB = na.omit(mergedB)
dim(mergedB)
```
##3.	In order to address our missing values for each column, it was important to omit missing values, duplicates and NAs. 

###Plot a chart after finding the meadians of each variable
```{r}
#4)	Computing the median Alcohol Content (ABV) and International Bitterness Unit (IBU) for each state. Plot a bar chart to compare.
#This gives us the median for alcohol content.
medianABV = median(mergedB$ABV)

#Getting the median of bitterness is given by this code. 
medianIBU = median(mergedB$IBU)

medFrame = mergedB %>% group_by(State) %>% summarise(medianABV = median(ABV), medianIBU = median(IBU))

#Plotting via bar charts, shows the Median ABV per State.
ggplot(medFrame, aes(x = State, y = medianABV)) + geom_bar(stat = 'identity') + labs(title = "Median ABV per State", y = "Median ABV")

#This will give us the Median IBU per State. 
ggplot(medFrame, aes(x = State, y = medianIBU), color = "yellow") + geom_bar(stat = 'identity') + labs(title = "Median IBU per State", y = "Median IBU")
```
###Arrange rows to find the max of each variable
```{r}
#5) Which state has the maximum alcoholic (ABV) beer? 
##Arranging the rows in descending order with reference to ABV.
arrange(medFrame, desc(medianIBU))

#Which state has the most bitter (IBU) beer?
#Arranging the rows in descending order with reference to IBU.
arrange(medFrame, desc(medianABV))

```
##The state with the maximum ABV and IBU happens to be Maine. 
###Histogram
```{r}
#6) Comment on the summary statistics and distribution of the ABV variable.
#Most beers average around 0.052. Looking at the overall distribution, the data 
#is right skewed. Our max for ABV is at 0.12500 and min is 0.02700.
ggplot(mergedB, aes(x = ABV)) + geom_histogram() + labs(title = "ABV Distribution", y = "Count")

```
##For the summary of the statistics and distribution of the ABV variable happens to be around 0.052. The data is right skewed. 

###Scatterplot
```{r}
#7) The relationship between the bitterness of the beer and its alcoholic content.
#Using a scatter plot we look into their relationship.
#The relationship between ABV and IBU appears to be linear and positive. There 
#are many outliers (noise) from what can be seen. 

ggplot(mergedB, aes(x = ABV, y = IBU)) + geom_point(size = 0.1) + geom_jitter() + stat_smooth(method = 'lm') + labs(title = "ABV vs. IBU")
```
##The relationship between the bitterness of the beer and its alcoholic content appears to be linear and positive. Thought there are many outliers that are represented, the general  observation is that the data are climbing. That is to say that the data are positively growing. 

###For Loop
```{r}
#8) Investigating the difference between IPAs and Ale with respect to IBU and ABV. 

#This guarantees the same values are produced.  
set.seed(7)

#Using grepl to find rows that contain IPA and Ale within the Style column. 
mergedBall= mergedB %>% filter(grepl("IPA|Ale", Style))

#Grabbing and then categorizing all Ales and IPAs through each row within the Sytle Column. 
for (i in 1:length(mergedBall$Style))
  {
  if (grepl("IPA", mergedBall$Style[i]))
    {
    mergedBall$group[i]= "IPA"
  } else{
    mergedBall$group[i] = "Ale"
  }
}

mergedBall
```
###Train the model
```{r}
#Creating the training and testing for IBU and ABV. Training the model. 
trainIndices = sample(1:dim(mergedBall)[1], round(0.75* dim(mergedBall)[1]))
train = mergedBall[trainIndices,]
test = mergedBall[-trainIndices,]

#Seeing visually how Ales and IPAs interact with a scatter plot.  
mergedBall %>% ggplot(aes(x = IBU, y = ABV, color = group)) + geom_point() + geom_jitter()

```

###For loop
```{r}

#This code chuck will loop through each iteration in order to find the best accuracy 
#for the k value. Once the best value is found it will be used in the next code chunk. 
iterations = 500
numks = 30
masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(30), k = numeric(30))
  trainIndices = sample(1:dim(mergedBall)[1], round(0.75* dim(mergedBall)[1]))
  train = mergedBall[trainIndices,]
  test = mergedBall[-trainIndices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(4,5)], test[,c(4,5)], train$group, prob = TRUE, k =3)
    table(classifications, test$group)
    cm = confusionMatrix(table(classifications, test$group))
    masterAcc[j,i] = cm$overall[1]
  }
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
```
###KNN
```{r}
#k = 23 is the max k value that gives us the best accuracy for the model 
classifications = knn(train[,c(4,5)], test[,c(4,5)], train$Style, prob = TRUE, k =23)
#table(classifications, test$Style)
#cm = confusionMatrix(table(classifications, test$Style))
#cm
# The model reports an Accuracy of 84.75% 
```

###Prediction test
```{r}

#Predicting to see if our data point will be an IPA or Ale using a low ABV and low IBU values
df_test = data.frame(ABV = 0.1 , IBU = 10 )

classifications = knn(train[,c(4,5)], df_test, train$group, prob = TRUE, k =3)
classifications 
#our model shows that the probably of it being an Ale as a 66% Accuracy 


#Using higher ABV and IBU values to test to see if those values give us an IPA
df_test = data.frame(ABV = 0.31 , IBU = 50 )

classifications = knn(train[,c(4,5)], df_test, train$group, prob = TRUE, k =3)
classifications 

#IPA has a higher average for ABV and IBU whilst Ale appears to have a lower average for ABV and IBU.
#This statistically reinforces the observation of IPA and Ales traits and/or relationship. 
mergedBall %>% group_by(group) %>% summarise(meanABV = mean(ABV), meanIBU = mean(IBU))

```
##The model predicts that it will be an IPA with a prob of 66%.It would seem that higher the IBU and ABV are, the higher the indicator is of a data point to be an IPA. 

###Filter and plot
```{r}
#9 Looking for eligible states for appropriate target market using ABV. If you want to make an IPA here are the states Budweiser should invest in. 
mergedBall %>% group_by(State) %>% summarise(meanABV = mean(ABV)) %>% arrange(desc(meanABV)) %>% 
  ggplot(aes(x = reorder(State, -meanABV), y = meanABV)) + geom_bar(stat = "identity") + 
  labs(x = "States", y = "Meadian ABV", title = "Meadian ABV by State")
```
##We should look for eligible states for appropriate target market by using ABV. If you want to make an IPA here are the states Budweiser should invest in. These are the top 5 states #1 " CT"  0.0745 #2 " NV"  0.0695 #3 " DC"  0.069  #4 " CO"  0.0674 #5 " MD"  0.0668.

###Filter and plot
```{r}
#If Budweiser wants to market or make Ale, here are the states Budweiser should look into. 
mergedBall %>% group_by(State) %>% summarise(meanIBU = mean(IBU)) %>% arrange(desc(meanIBU)) %>% 
  ggplot(aes(x = reorder(State, -meanIBU), y = meanIBU)) + geom_bar(stat = "identity") + 
  labs(x = "States", y = "Meadian IBU", title = "Meadian IBU by State")

```
##If you want to market or make Ale, here are the states Budweiser should look into. These are the top 5 states# 1 " AR"  0.04   #2 " KS"  0.0528 #3 " ND"  0.054  #4 " WY" (Wyoming) 0.0547 #5 " MO"  Missouri 0.0549 
